---
title: "InvCor"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("D:/UMCG/Genetica/Projects/Depict2Pgs")


library(heatmap3)
library(RColorBrewer)

```

## Genes in arm: 100 number of permutation to determine correlation: 10


```{r }
numberRandomGwas=10
genesInArm=100
data <- matrix(rnorm(genesInArm*numberRandomGwas), ncol = genesInArm)
corData <- cor(data)
heatmap3(corData, scale= "none", Rowv = NA, Colv = NA, col = colorRampPalette( brewer.pal(9, "YlOrBr"))(100))
det(corData)
```

## Genes in arm: 100 number of permutation to determine correlation: 100

```{r }
numberRandomGwas=100
genesInArm=100
data <- matrix(rnorm(genesInArm*numberRandomGwas), ncol = genesInArm)
corData <- cor(data)
heatmap3(corData, scale= "none", Rowv = NA, Colv = NA, col = colorRampPalette( brewer.pal(9, "YlOrBr"))(100))
det(corData)
```


## Genes in arm: 100 number of permutation to determine correlation: 1000

```{r }
numberRandomGwas=1000
genesInArm=100
data <- matrix(rnorm(genesInArm*numberRandomGwas), ncol = genesInArm)
corData <- cor(data)
heatmap3(corData, scale= "none", Rowv = NA, Colv = NA, col = colorRampPalette( brewer.pal(9, "YlOrBr"))(100))
det(corData)
```


## Genes in arm: 1000 number of permutation to determine correlation: 1000

```{r }
numberRandomGwas=1000
genesInArm=1000
data <- matrix(rnorm(genesInArm*numberRandomGwas), ncol = genesInArm)
corData <- cor(data)
heatmap3(corData, scale= "none", Rowv = NA, Colv = NA, col = colorRampPalette( brewer.pal(9, "YlOrBr"))(100))
det(corData)
```

## Genes in arm: 1000 number of permutation to determine correlation: 10000

```{r }
numberRandomGwas=10000
genesInArm=1000
data <- matrix(rnorm(genesInArm*numberRandomGwas), ncol = genesInArm)
corData <- cor(data)
heatmap3(corData, scale= "none", Rowv = NA, Colv = NA, col = colorRampPalette( brewer.pal(9, "YlOrBr"))(100))
det(corData)
```

## Genes in arm: 1000 number of permutation to determine correlation: 100000

```{r }
numberRandomGwas=100000
genesInArm=1000
data <- matrix(rnorm(genesInArm*numberRandomGwas), ncol = genesInArm)
corData <- cor(data)
heatmap3(corData, scale= "none", Rowv = NA, Colv = NA, col = colorRampPalette( brewer.pal(9, "YlOrBr"))(100))
det(corData)
```

## Identity matrix has det of 1
```{r}
genesInArm=1000
corData <- diag(x=1, nrow = genesInArm)
heatmap3(corData, scale= "none", Rowv = NA, Colv = NA, col = colorRampPalette( brewer.pal(9, "YlOrBr"))(100))
det(corData)

```


